\clearpage
\phantomsection
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENTS}
\begin{center}
ACKNOWLEDGMENTS
\end{center}

\begin{doublespacing}

	I would like to begin by expressing my deepest gratitude to my thesis
advisor, Dr.\ Oscar Morales-Ponce, whose mentorship has been one of the
most defining parts of my time at California State University, Long
Beach.
He met me as a freshman who did not yet know how to use a terminal, and
he taught me with patience, bore with me through every stumble, and
watched me grow into the researcher and soon-to-be graduate I am today.
I cannot picture my CSULB years without his guidance, his generosity
with his time, and his unwavering belief in what I could become.

	I am also deeply grateful to the broader CSULB community that
surrounded and sustained me.
To the international student advisors, the faculty, and the many staff
members who took the time to answer an email or spare a minute in the
hallway: every one of those small moments added up to something I will
carry with me far beyond graduation.
I owe a particular debt of gratitude to the entire BESST community, who were there for me through thin and thick. I wish I could write a second thesis on how their support found its
way into every word of this one. 

	Last, and most deeply, I want to thank my family.
To my aunt, who opened her home to me when I first arrived and became
not only a roof over my head but my backbone, my confidante, and in
every meaningful sense my family here in the United States: I am
endlessly grateful.
To my best friend, who has been both a steadfast companion and a big
sister to lean on through every high and low: thank you for being
present in the ways that mattered most.
To my grandparents, who remained a constant source of encouragement
and motivation at every step of this journey: your love reached me
across every distance.

	To my brother, who supported every path I chose without hesitation
and whose creative mind I have always looked up to: you have shaped
the way I see the world and approach every problem.
To my father, whose support never wavered even through the hardest
financial times: your steadiness gave me the stability to keep going.
And to my mother, who made the most quietly courageous decision of all
by letting her heart break, and mine, so that I could leave Iran and
build this life: every page of this thesis belongs to you as much as
it belongs to me.
I love you.

\end{doublespacing}
\clearpage
