\chapter{CONCLUSIONS AND FUTURE WORK}\label{ch:conclusion}

\begin{doublespacing}

\section{Summary of Findings}

This thesis has presented a unified framework for analyzing multi-robot
search and rescue across four algorithmic models that vary energy
capacity and communication assumptions. Theoretical upper bounds on
the expected finder competitive ratio were derived for each model and
validated by Monte Carlo simulation over 500 trials per configuration.

The headline contribution is the identification of a bid function
$p_i / d^2$ that outperforms the conventional $p_i / d$ under energy
constraints, justified by a \emph{cost-foreclosure} argument: a
distant anchor site not only costs proportionally more energy but
also shrinks the remaining chain-extension budget by the same amount,
so the effective marginal cost of distance is superlinear. The result
is practical in two senses. It is a one-line change in the bid
function of any auction-based multi-robot search system, and the
argument behind it is generic---it applies wherever a fixed resource
budget must be allocated between a committed first action and a
sequence of optional follow-ups.

Three further experimental results strengthen and extend the main
findings. First, the $p/d^2$ advantage over $p/d$ persists across
prior-misspecification noise levels up to $\sigma = 1.0$; the
absolute margin in M4*'s favour \emph{grows} with noise, from $0.13$
FCR units on clean priors to $0.49$ units at moderate noise. Second,
the $p/d^2$ advantage also persists under three non-Euclidean cost
models (Manhattan distance, heterogeneous robot speeds, obstacle
detour penalties); the foreclosure argument is therefore robust to
the specific metric used. Third, the thesis identifies a named
finding called the \emph{Target-Delay Anomaly}: post-processing a
greedy $p/d^2$ chain with a standard 2-opt tour-length optimizer
\emph{worsens} finder FCR by $5.3$ percent despite shortening the
tour, because 2-opt reorders sites by geometry and so dilutes the
probability-weighted prefix that a finder-centric metric rewards.
This is a non-obvious conclusion: the standard robotics pipeline of
``build a tour, then polish it'' is not neutral but actively
harmful under Bayesian priors and finder-centric objectives.

Four headline findings plus three extensions emerged, each
corresponding to one of the research hypotheses stated in
Chapter~III.

\emph{Coordination gain (H1).}
Auction-based coordination yields a $6.6\times$ improvement in
finder competitive ratio over random uncoordinated search (M1 FCR $= 18.74$
vs.\ M2 FCR $= 2.83$ over 5,000 trials), confirming
that the communication mode---not the fleet size---is the dominant
determinant of search efficiency, consistent with the theoretical
prediction of \citet{Chrobak2015}.

\emph{Mechanism matters (H2).}
Probability-aware distributed auctions (M3 FCR $= 6.32$) outperform a centralized
distance-optimal Hungarian baseline (H$_d$ FCR $= 6.88$) by $8.1$ percent.
A probability-weighted Hungarian ablation H$_{p/d}$ (FCR $= 7.38$)
performs \emph{worse} than H$_d$, demonstrating that the SSI auction's
iterative site removal---not the bid function alone---is the essential
mechanism. Giving the Hungarian solver probability-weighted costs without
changing its one-shot centralized assignment structure introduces a
clustering pathology that is absent in the sequential SSI auction.

\emph{Energy recovery (H3).}
Multi-node sorties recover $90$ percent of the gap between the
energy-constrained single-sortie algorithm and the
unconstrained-energy auction baseline. A single greedy chain extension
per sortie visits $\sim\!5.3$ sites on average at the default setting,
cutting the iteration count from $4.81$ to $1.38$.

\emph{Bid function (H4).}
A bid function $p_i / d^2$ that penalizes distance quadratically
improves on the standard $p_i / d$ bid (M4* FCR $= 3.11$ vs.\ M4 FCR $= 3.16$,
a $1.5$ percent directional improvement) under energy
constraints. The result is justified by a cost-foreclosure argument:
distant anchor sites not only cost more energy but also shorten the
chain of future extensions, so the effective cost of distance is
superlinear.

All four theoretical bounds held empirically across every trial. The
bounds on Models~1 and~2 are relatively tight; the Model~4 bound is
loose because the analytical estimate of sites per sortie
underestimates what the greedy chain actually achieves.

\section{Future Work}

Several directions for extending this work are natural.

\textbf{Alternative objectives.} Finder competitive ratio measures
only the finding robot's distance; it says nothing about the other
robots' energy expenditure. Minimax workload balancing, total-fleet
distance, or time-to-last-cleared-site are alternative objectives that
would favor different bid functions. A systematic study of the
trade-off among these objectives on a common benchmark is an open
problem. The sensitivity analysis framework of \citet{AutomaticaAuctions2023}
provides formal machinery to characterize how the optimal bid exponent
shifts as the objective changes.

\textbf{Tighter Model~4 bound.} The looseness of the Model~4 bound
comes from the pessimistic estimate $S = 1 + \lfloor (E -
2\,d_{\mathrm{avg}}) / d_{\mathrm{hop}} \rfloor$. A more refined
analysis would compute the expected length of a greedy chain under the
prior distribution, not the worst-case length under a uniform
distribution. Because the greedy chain preferentially selects sites
closer than $d_{\mathrm{hop}}$, the expected $S$ is better described
by the expected minimum of a random sample from the inter-site
distance distribution---a quantity that can be computed in closed form
for uniform site placements. This tighter bound would bring the
analytical estimate from the current factor-of-$4.5$ gap down to
within a factor of two of the empirical mean.

\textbf{Fleet-size scaling (proved).} The robot sweep
(Chapter~V, Figure~5) is explained by the fleet-size scaling results
proved in Section~\ref{sec:scaling} (Chapter~IV).
In the experimentally relevant regime $1 \leq R \leq n/2$, both M1 and M2
achieve $\mathbb{E}[\mathrm{FCR}] = \Theta(1/\sqrt{R})$, with M2's
constant factor $\Theta(1/\sqrt{n})$ times smaller than M1's.
The key lemma is $\mathbb{E}[D_{\mathrm{opt}}] = L/(2\sqrt{R}) + O(L/R)$,
proved via the survival function of the minimum of $R$ i.i.d.\ uniform
base-to-target distances.
At $n = 30$, the coordination advantage is predicted to be
$\approx 1/\sqrt{30} \approx 5.5\times$, consistent with the
observed $6.6\times$ gap (within factor-of-mean approximation corrections).
This rigorously confirms that communication mechanism, not fleet size,
is the primary design lever in multi-robot search.

\textbf{Comparison with learning-based approaches.} Recent DRL
methods for multi-robot search~\citep{KimKim2025} optimize richer
posterior objectives (Fisher information, mutual information) and
can exploit complex spatial patterns that closed-form bid functions
cannot represent. A head-to-head evaluation of M4* against a trained
DRL policy on the same instance distribution would quantify precisely
how much is lost by restricting to analytical bids---and, conversely,
how much is gained by having provable bounds. The Target-Delay
Anomaly (Chapter~V) already suggests one dimension on which DRL
policies trained on average-case path length would lose to M4*: they
would likely learn to minimize tour length rather than
probability-weighted prefix, and so suffer from the same reordering
effect that made 2-opt harmful.

\textbf{Relaxing assumptions.} Perfect sensors, flat terrain, and
uniform energy budgets are idealizations. Detection probability less
than one, shortest-path distance over an obstacle graph, and per-robot
energy heterogeneity are realistic generalizations. The auction
mechanism adapts to each without modification; what changes is the
structure of the bounds. For imperfect sensors with detection
probability $p_{\mathrm{det}} < 1$, Lemma~1 generalizes by replacing
$R$ with $R \cdot p_{\mathrm{det}}$, and the bid update rule becomes
$p_i \leftarrow (1 - p_{\mathrm{det}}) p_i$ rather than zero.

\textbf{Communication degradation.} \citet{OtteKuhlman2020} showed
that different auction protocols degrade at very different rates under
unreliable communication. The performance of M4* under dropped or
delayed radio links---a condition that is common in real rubble-pile
deployments---is unknown. A natural extension is to evaluate M4* under
the communication models of \citeauthor{OtteKuhlman2020} and derive
degraded-communication FCR bounds.

\textbf{Physical deployment.} \citet{Coad2024SPROUT} demonstrated
that vine-robot deployment in USAR scenarios currently operates in
single-robot mode; multi-robot coordination remains an open
deployment challenge. Porting M4* to a small team of UAVs or ground
robots and measuring FCR on physical instances---including sensor
noise, communication loss, obstacle avoidance, and motor control
variability---would provide the strongest validation of the framework
and is the most operationally impactful next step.

\end{doublespacing}
