\chapter{INTRODUCTION}\label{ch:intro}

\begin{doublespacing}

\section{A Rescue Scenario}

Imagine a five-story apartment building has just collapsed after a
major earthquake. The debris covers an area of roughly half a city
block. Somewhere inside, a person is trapped in a void pocket under a
slab of concrete. The fire department arrives, but the structure is
unstable and they cannot send people in safely. Instead they deploy
three small autonomous drones---rugged, battery-powered, each one
about the size of a shoebox. Each drone is launched from a truck
parked on the street, and each drone's battery will last for roughly
fifteen minutes of flight time before it has to return to the truck
to swap batteries.

From previous search-and-rescue experience, the rescuers already
have a rough idea of where in the rubble survivors are most likely to
be found: near stairwells, around load-bearing walls, under furniture
that might have created a void pocket. Collectively the rescuers
identify about thirty such ``candidate sites.'' Each site has a prior
probability associated with it---some sites are very likely to hold
the trapped person (near a bathroom, say, where someone might have
taken shelter), and some are unlikely but non-zero (under the
collapsed roof).

The three drones need to check these thirty sites, one by one, until
they find the trapped person. The question this thesis asks is
simple to state: \emph{what strategy should the drones use to find
the person as fast as possible?} ``Fast'' here means short total
travel distance, because distance is what drains the battery and
delays the rescue.

The question sounds simple, but it hides four separate subproblems
that interact in surprising ways:

\begin{enumerate}
\item \emph{Which drone should check which site?} If the three
drones all rush toward the most probable site, two of them waste a
trip. If they don't coordinate at all, they will sometimes end up
checking the same site twice.

\item \emph{How should a drone decide where to go next?} The obvious
answer is ``the nearest site,'' but the nearest site might have a
very low probability of actually holding the survivor. A site a
little farther away with a much higher probability is usually a
better bet.

\item \emph{How should a drone use its limited battery?} A drone
that flies to a site, checks it, and flies back to the truck has
used up battery mostly on travel, not on searching. If instead it
chains together several sites in one trip before returning, it
searches far more effectively. But chaining too many sites risks
running out of battery mid-flight.

\item \emph{What if communication between the drones drops out?} In
a real disaster site, radios fail. If the drones cannot talk to each
other while in flight, coordination has to happen back at the truck,
which changes everything about how decisions get made.
\end{enumerate}

This thesis answers all four questions, using a mix of formal
mathematical analysis and computer simulation.

\section{What This Thesis Shows, in One Paragraph}

The main result is the following. Of all the factors that affect how
fast the drones find the survivor, the single most important one is
\emph{how they decide which site to go to next}. The conventional
choice is to rank sites by ``probability divided by distance,''
meaning a site with twice the probability at twice the distance
ranks the same as one with half the probability right next door.
This thesis shows that under battery constraints the right rule is
instead ``probability divided by distance \emph{squared}.'' The
quadratic penalty on distance makes distant sites look twice as bad
as the standard rule would suggest, and this is exactly correct when
battery is scarce, because going to a far site costs you not just
that trip's battery but also the opportunity to chain several
closer sites afterward. Empirically, the new bid function improves
average search time by about four percent on top of an already
well-optimized baseline---a small number, but one that comes for
free from a one-line change to the algorithm.

Several other findings fall out of the same analysis. Coordination
via an auction mechanism improves search time by 6.6$\times$ over
robots choosing independently---so communication, not fleet size, is
the main lever. Allowing drones to chain multiple sites per flight
recovers 92\% of the performance gap between battery-constrained and
battery-unlimited operation. These results are all formally backed
by upper bounds on the worst-case search distance that are proved in
Chapter~IV and verified over 500 simulation trials per configuration
in Chapter~V.

\section{Key Terms at First Encounter}

Before diving in, a few technical terms appear throughout the thesis
and deserve a plain-language definition. They are introduced
informally here and formally in Chapter~III.

\begin{description}[leftmargin=2em,style=nextline]
\item[Robot.] An autonomous battery-powered agent that can travel to
a point in space, check whether the target is there, and return. In
the running example, ``robot'' means a drone, but the same algorithms
work for any kind of mobile agent: ground robots, underwater vehicles,
or even teams of humans.

\item[Site.] One of the $n$ candidate locations the robots might
check. Each site has a fixed position in a two-dimensional search
area and a prior probability of containing the target.

\item[Prior probability.] How likely a site is to contain the target,
before any searching happens. Priors come from domain knowledge---a
structural engineer's assessment of which voids are survivable, a
search dog's indication, etc. Priors always sum to one across all
sites, because the target is somewhere.

\item[Target.] The object the robots are trying to find. In the
running example, a survivor; in other applications, a lost hiker, a
chemical leak, or a piece of evidence.

\item[Base.] The location each robot returns to for recharging. In
the running example, the fire truck. Different robots can have
different bases; the algorithm does not assume they all share one.

\item[Sortie.] A single round trip by one robot: base $\rightarrow$
one or more sites $\rightarrow$ base. A sortie is limited by the
robot's battery energy budget $E$.

\item[Auction.] A formal protocol the robots use to decide who goes
where. Each robot computes a ``bid'' for each site---a numerical
score saying how attractive that site is to that robot---and the
highest bidder wins the site. Auctions in this thesis are always
\emph{sequential single-item} (SSI) auctions: one site is auctioned
at a time, the winner is assigned, and the process repeats until all
robots have an assignment.

\item[Bid function.] The mathematical formula that turns a (site
probability, site distance) pair into a bid. The thesis studies
several: $p$ alone (only consider probability), $1/d$ alone (only
consider distance), $p/d$ (the classical choice), $p/d^2$ (the one
this thesis recommends), and a few others.

\item[Finder Competitive Ratio (FCR).] The main performance metric
of the thesis. It is the total distance traveled by the robot that
actually finds the target, divided by the straight-line distance
from the nearest base to the target. An FCR of 1.0 means the robot
flew directly to the survivor without ever going the wrong way; an
FCR of 10.0 means the robot wandered ten times as far as strictly
necessary. Lower is better. FCR is formally defined in Chapter~III
and illustrated with a worked example in Figure~\ref{fig:fcr-explained}.
\end{description}

\section{Why This Problem is Not Already Solved}

A reader with some background might reasonably ask: surely search
robots have been studied for decades, and surely someone has solved
this already? The answer is that the problem has been studied from
three different angles, each in isolation, and no prior work puts all
three together. Chapter~II explains the three strands in detail; in
brief:

\begin{itemize}
\item \textbf{Competitive search theory}, starting with the cow-path
problem~\citep{BaezaYates1993}, produces elegant worst-case
guarantees, but the geometries it studies are one-dimensional
(``walk along a line looking for a cow'') and the battery is
unlimited.
\item \textbf{Market-based multi-robot coordination} (auctions)
scales to large teams and works well in practice, but the standard
treatment assumes robots can fly as far as they want.
\item \textbf{Energy-constrained exploration} studies fuel-limited
robots, but typically without formal competitive-ratio guarantees
and without auction-based coordination.
\end{itemize}

Putting all three together---finite battery, auction coordination,
and a competitive-ratio performance guarantee, in a two-dimensional
arena---is the contribution of this thesis.

\section{Motivation from the Field}

The problem is not hypothetical. Urban search-and-rescue (USAR)
robotics has matured over the last two decades from tele-operated
single-unit deployments (one human piloting one robot over a tether)
toward teams of semi-autonomous robots. Field reports from
deployments at the World Trade Center collapse~\citep{Casper2003},
from the Fukushima Daiichi nuclear accident response, and from
several recent earthquake responses consistently identify two
practical pain points.

The first is \emph{battery}. Real rescue robots exhaust their
batteries before finishing their search, and the time spent swapping
batteries is time survivors do not have. Algorithms that waste any
portion of a battery are therefore not merely inefficient---they can
cost lives.

The second is \emph{duplicated effort}. When two robots cover the
same rubble without coordination (typically because a radio link
dropped), every minute of one robot's work was wasted from the
team's perspective. Poor coordination causes robots to check the
same locations and to burn their finite batteries before checking
the most likely ones.

The two pain points are related. Under strict energy budgets, each
wasted sortie---each excursion that checks a site of low prior
probability or revisits a site that a teammate already cleared---is
a larger fraction of the total searchable volume. Coordination
quality therefore matters \emph{more} as energy grows scarcer, not
less, which inverts the naive intuition that a ``self-sufficient''
robot with enough battery can simply brute-force the problem. Giving
each robot a bigger battery does not remove the need for good
coordination; it just shifts the failure mode.

\section{Scope and Contributions}

The scope of this thesis is algorithmic and simulation-based. Four
algorithmic models that span the energy--communication design space
are formulated, closed-form upper bounds on their expected Finder
Competitive Ratio (FCR) are derived, and the bounds are validated
through 500 Monte Carlo trials per configuration on identical
instances, with parameter sweeps over the energy budget $E$ and the
fleet size $R$. Physical robot deployment is not in scope and is
left as future work, discussed in Chapter~VI.

The specific contributions of this thesis are as follows:

\begin{enumerate}
\item A formalization of four search models---M1 (random), M2
(unconstrained-energy auction), M3 (single-sortie auction), and
M4/M4* (multi-sortie auction)---that systematically vary the two
operational dimensions of energy and communication, together with a
closed-form expected-FCR upper bound for each.

\item An ablation on the Hungarian centralized assignment baseline
that isolates two previously conflated effects: the value of using
probability-aware bids versus the value of centralized allocation.
The new experimental result shows that probability-awareness is the
dominant factor.

\item Identification of a bid function $p_i / d^2$ that
\emph{empirically outperforms} the conventional $p_i / d$ bid under
energy constraints, together with a \emph{cost-foreclosure} argument
that justifies why the distance penalty should be superlinear: a
distant anchor site not only costs proportionally more energy to
reach, it also shrinks the remaining energy budget available for
chain extensions by the same amount. This is the primary
contribution of the thesis.

\item Validation of all four upper bounds across 500 Monte Carlo
trials, and parameter sweeps over $E \in [8, 30]$ and $R \in [1, 6]$
showing the stability of the model ranking and the
diminishing-returns behavior of fleet size.
\end{enumerate}

\section{How to Read This Thesis}

The chapters build on each other but can be read at different
depths. Readers without a multi-robot systems background may wish to
start with Chapter~II's literature primers, skim the formal
Chapter~III only for the definitions of FCR and the sortie model,
and focus on Chapter~V's results, revisiting Chapter~IV only for
specific theorems as needed. Readers with a robotics background may
prefer the reverse order: the formal problem and algorithms in
Chapters~III--IV, then the experimental validation in Chapter~V.

Chapter~II reviews the related work in competitive search theory,
auction-based coordination, and probabilistic and energy-constrained
search, and positions the present contribution relative to each.
Chapter~III states the formal problem, defines the Finder
Competitive Ratio, and introduces the sortie model and the geometric
constants used in the bounds. Chapter~IV presents the four
algorithmic models and the Hungarian baselines, together with the
theoretical upper bounds on the expected FCR. Chapter~V describes
the experimental setup and reports the main comparison results, the
bounds verification, and the parameter sweeps, and discusses the
filtering thresholds and the simplifying assumptions. Chapter~VI
concludes with a summary of findings and a discussion of future
work.

\end{doublespacing}
